---
title: "Stat 210 Vectors"
output: html_notebook
---

There are two types of vectors, Atomic and Lists

Atomic consist of subtypes: logical, integer, double, character, complex, and raw

Lists Can contain other lists. Lists can hold multiple types within

NULL - absence

type() will return the type of a vector
typeof()
length() will return the length/size
When using floats in R use dplyr::near() to evaluate instead of ==

Additional meta-data can be within a vector. 3 important types of augmented vectors Factors, Dates, Data Frames & Tibbles

Place an L after a number to make it an integer

Special values NA, NaN, Inf and -Inf These also should not use == but use instead
is.finite()
is.infinite()
is.na()
is.nan()

Strings simply use pointers to the same object to cut down on memory usage
Rare missing value data types 
NA            logical
NA_integer_
NA_real_      double
NA_character_ Char



```{r}
library(tidyverse)

#logical
c(TRUE, TRUE, FALSE, NA)

#Colon notation generates from 1 to 10. Performs modulus on each item in the list
# If the modulus result equals 0 .... Implicit statement?
1:10 %% 2 == 0


```



Exercises
```{r}
x <- NaN
is.finite(x)
!is.infinite(x)

```
These two things aren't evaluating the same thing otherwise you would expect the ! to make them return the same values.
Missing values are not considered finite, while is.infinite considers just Inf and -Inf to be infinite

```{r}
dplyr::near
```

The function works by checking if the difference in absolute value is less than a number

3. A integer vector can take 11 possible int values 0-9 & NA upto 2^32
   A double can represent upto 2^64 with -Inf, Inf,NA, and Nan being part of those values.
4. floor (round down), ceiling (round up), round (round up at .5 down at .4), and.... trunc() which rounds towards zero
5. parse_logical (replace second part with integer and number)

```{r}
x <- sample(20, 100, replace = TRUE)
y <- x > 10 
```


Useful functions to check data type
is_logical()
is_integer()
is_double()
is_numeric()
is_character()
is_atomic()
is_list()
is_vector()

scalar version checks that the length is one is_scalar_logical()

Tidyverse doesn't like to automatically 'recycle' smaller size vectors to match larger ones for operations. if you want to do it anyways use 
tibble(x = 1:4, y = rep(1:2, 2))

you can name vectors eg c(x = 1, y = 2, z = 3)
you can name them after the fact too set_names(1:3, c("a","b","c"))
subsetting within a vector is like an index in an array x[a] and will return the value at a
x[c(3,2,1)] yields 'three', 'two', 'one'
Negative values counter intuitively drop values out of the output. don't mix + and -
x[c(-1,-3,-5)]
x[] blank means all
In R you can use names to access vector values e.g.
x["xyz", "def"]
5 2


[[ extract a single item, not meta like attached names

1. 
```{r}
is.na(x) #returns true is NA false if not
mean(is.na(x))#adds up the 0's and 1s and divides by the total, tells you how much is na

```

2. is vector checks for a vector, and names, but if it has anything else it's not considered a vector. is atomic checks for one of the basic types
3. purr simply has more ways it allows you to set names
4. 
```{r}
singled_out <- function(x){
  x[[length(x)]]
}
 singled_out(c(1,2,3,4,5,6,7,8,9)) 
```

```{r}
No0ddsAllowed <- function(x){
  x[x %% 2 == 0]
}
 No0ddsAllowed(c(1,2,3,4,5,6,7,8,9)) 
```
```{r}
bully_last_pick <- function(x){
  x[-length(x)]
}
bully_last_pick(c(1,2,3,4,5,6,7,8,9)) 
```
```{r}
getting_even <- function(x){
  x[is.na(x) %% 2 == 0]
}
getting_even(c(1,2,3,4,5,6,7,8,9,NA)) 
```
5. One includes Na the other does not
6. it returns Na




You can create a list with list()
str() shows what STRucture is in a list
Lists can have multiple object types within it
list(list(1,2),TRUE, "a")

a <- list(a = 1:3, b= " astring", c=pi d = list(-1,-5))

to access elements within a list
[a] - returns list object with selected elements within it (pepper shaker with packets within it)
[[a]] - returns just the element within the list
[[a]][[x]] - returns the x element within the element a of the list
$ is used to access and single element a$a is equivalent to a[["a]]

2. The major difference between a tibble and list is that tibbles must have the same number of rows, while a list can vary. A list can contain a data frame or tibble.


attributes
Vectors can contain additional meta data called attributes.
set attributes with attr()
see them with attributes()
important for adding names, Dimensions, or Class

x <- 1:10
attr(x, "greeting")
#> NULL
attr(x, "greeting") <- "Hi!"
attr(x, "farewell") <- "Bye!"
attributes(x)
#> $greeting
#> [1] "Hi!"
#> 
#> $farewell
#> [1] "Bye!"


R can be used with object oriented programming.
You have generic methods such as as.date() which depending on the arguement type that is passed to it
will be handled by a specific methoded contained within the (I am guessing) class.
methods("as.date") will display all the specific methods. getS3method() shows the implementation

1.
```{r}
hms::hms(3600)
j <-hms::hms(3600)

class(j)
typeof(j)
```
It returns a time, that is an object of difftime, which holds a double

2. It generates an error if different size vectors are used.
3. So long as the list is the correct length it is okay...even if there are different types involved.
